---
title: "DAS GROUP 8 PROJECT "
author: "Constantinos Anastasiou, Lin Lin, Yuang Tian, Yutong WU, Shengyuan Xia"
number-sections: true
format: 
  html:
    embed-resources: true
    code-tools: true
  pdf: default
editor_options: 
  chunk_output_type: console
execute:
  echo: false
  eval: true
  warning: false
  message: false
---

```{r}
#| echo: false
library(ggplot2)
library(tidyverse)
library(gt)
library(patchwork)
library(gridExtra)
library(moderndive)
library(knitr)
```

# Introduction {#sec-Intro}

# Exploratory Data Analysis {#sec-EDA}

## **Data Summary**

```{r}
library(tidyverse)
library(gt)
library(corrplot)
mov <- read_csv("C:/Users/admin/Desktop/dataset08.csv")

# solve NA (delete the NA rows)
mov_1 <- na.omit(mov)

# 1. data summary
library(gt)
summary(mov_1)
mov_1 %>%
  select(-film_id) %>%
  select(where(is.numeric)) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value") %>%
  group_by(Variable) %>%
  summarise(
    'Mean' = mean(Value),
    'Median' = median(Value),
    'St.Dev' = sd(Value),
    'Min' = min(Value),
    'Max' = max(Value),
    'IQR' = IQR(Value),
  ) %>%
  gt() %>%
  fmt_number(decimals = 2) %>%
  cols_label(
    Variable = "Variable",
    Mean = "Mean",
    Median = "Median",
    St.Dev = "Std. Dev",
    Min = "Minimum",
    Max = "Maximum",
    IQR = "IQR"
  )
```

## **Histograms for Numeric Variables**

```{r}
#univariate analysis
# histogram

ggplot(mov_1, aes(x = rating)) + 
  geom_histogram(binwidth = 0.5, fill = "#69b3a2", color = "#404040") +  # 使用自定义颜色
  labs(title = "Rating Distribution", x = "IMDB Rating", y = "Count")
```

1.The distribution of ratings shows a bimodal trend, with one peak occurring at about 3.8 and the other at about 7.6. This could mean that films are either getting above average ratings or higher ratings, with fewer films getting intermediate ratings.

3.Relatively few films received very low (below 2.5) or very high (close to 10) ratings.

4.The number of highly rated films seems to be greater than the number of lowly rated films, probably because IMDB users are more inclined to rate films they like.

## **Boxplots for Numeric Variables**

```{r}
# boxplot - budget
p2 <- ggplot(mov_1, aes(x = genre, y = rating, fill = genre)) +  # 在aes()中设置fill参数
  geom_boxplot() +
  scale_fill_brewer(palette = "Set3") +  # 使用Set3配色方案
  labs(x = "Genre", y = "Rating", title = "Rating by Genre")
print(p2)
```

## **Bar chart - genre**

```{r}
mov_1$rating_category <- ifelse(mov_1$rating >= 7, '7+', 'Below 7')
ggplot(mov_1, aes(x=genre, fill=rating_category)) +
  geom_bar() +
  scale_fill_manual(values=c('7+' = 'blue', 'Below 7' = 'skyblue')) +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title="Movie Genre Frequency by Rating Category", x="Genre", y="Count")
```

This bar chart shows the frequency distribution of different film genres. According to the chart: The Action and Drama genres have the highest number of films, with Drama having the highest number, followed closely by Action. Documentary and Comedy are relatively few in number.Animation, Romance and Short films are the least numerous, with Short films being the lowest.

## Scatter plots

```{r}
# Scatterplot - Budget & rating
ggplot(mov_1, aes(x=budget, y=rating)) +
  geom_point(aes(color=genre), alpha=0.5) + 
  scale_color_brewer(palette="Set2") +
  labs(title="Budget vs. Rating by Genre", x="Budget", y="IMDB Rating")
ggplot(mov_1, aes(x=length, y=rating)) +
  geom_point(aes(color=genre), alpha=0.5) + 
  scale_color_brewer(palette="Set2") +
  labs(title="Length vs. Rating by Genre", x="length", y="IMDB Rating")
ggplot(mov_1, aes(x=votes, y=rating)) +
  geom_point(aes(color=genre), alpha=0.5) + 
  scale_color_brewer(palette="Set2") +
  labs(title="Votes vs. Rating by Genre", x="votes", y="IMDB Rating")
ggplot(mov_1, aes(x=year, y=rating)) +
  geom_point(aes(color=genre), alpha=0.5) + 
  scale_color_brewer(palette="Set2") +
  labs(title="Year vs. Rating by Genre", x="year", y="IMDB Rating")

```

### Budget-rating relationship

The graph shows the distribution of different types of films (indicated by different colours) in terms of budget and IMDB rating. Overall, there is no obvious direct relationship between budget and IMDB rating, even for films with higher budgets, their ratings cover a wide range from low to high.


## Correlation coefficient matrix

```{r}
par(mfrow=c(1,1), mar=c(0,0,2,0))  
numerical_mov_1 <- mov_1 %>% select(rating, year, budget, votes, length) 
cor_matrix <- cor(numerical_mov_1, use="complete.obs") 
corrplot(cor_matrix, method="circle", addCoef.col = "black",  # 添加黑色的颜色键
         tl.cex = 0.7,            
         cl.cex = 0.7,            
         tl.srt = 0 )            
```

-   **Votes and Budget**: There seems to be a moderately strong positive correlation between votes and budget, suggesting that films with larger budgets might also receive more votes, possibly indicating higher viewer engagement or more extensive marketing campaigns.

-   **Votes and Length**: There appears to be a small to moderate positive correlation, which might suggest that longer films tend to receive more votes. This could reflect a trend that more substantial, possibly more serious films garner more engagement.

-   **Rating and Length**: There is a small negative correlation indicated by a light red color, suggesting that longer films might have slightly lower ratings, but the relationship seems weak.

-   **Rating and Budget**: There seems to be very little to no correlation between budget and rating, indicating that a larger budget does not necessarily correspond to higher ratings.

## Kruskal-Wallis Test

```{r}
# We want to test whether there is a significant difference in film ratings across genres
library(stats)
kruskal.test(rating ~ genre, data=mov_1) # Kruskal-Wallis test
```

According to the Kruskal-Wallis test above, this very small p-value suggests that we have enough evidence to reject this null hypothesis to conclude that there is a significant difference in the ratings of different types of films. In other words, the genre of a film is likely to have an effect on its IMDb rating.

# Formal Data Analysis {#sec-FDA}

```{r}
library(tidyverse)
library(moderndive)
library(gapminder)
library(sjPlot)
library(stats)
library(jtools)



#CLEANING DATA
data <- read.csv("dataset08.csv")
films <- na.omit(data)

```

We are fitting the following logistic regression model as seen below:

$$
g(p_i) = \log\left(\frac{p_i}{1-p_i}\right) = \mathbf{x}_i^\top \boldsymbol{\beta}
$$

where $p_i$ is the probability of getting a film rating higher than 7, and $\mathbf{x}_i^\top$ is the vector of covariates of the $i$th film.

```{r}
#WE CONVERT RATING TO BINARY VARIABLE
data <- read.csv("dataset08.csv")
films <- na.omit(data)
films$GREAT <- ifelse(films$rating > 7 , 1 ,0)
levels(films$GREAT) <- c("less rating than 7" , "greater rating than 7")

#FIT GLM
films$genre <- as.factor(films$genre)
glm <- glm(GREAT ~ year + length + budget + genre + votes , data = films , family= binomial(link = logit))
summary(glm)
#refined model without votes as it is not statistically significant
glm1 <- glm(GREAT ~ year + length + budget + genre    , data = films , family= binomial(link = logit))
```

## Interpretation of coefficients

Results indicate that a one year change (one unit change) in release multiplies the odds of a film receiving a rating greater than 7 by approximately exp(0.010238)= 1.01. Conversely, longer films tend to have slightly lower odds of receiving a rating greater than 7, with each additional minute multiplying the odds by approximately exp(-0.056869 ) = 0.94. Budget demonstrates a significant positive association, with each additional unit of budget (in \$1000000s) multiplying the odds of a film receiving a rating greater than 7 by approximately exp(0.509979) = 1.67. Positive voting from people pretty much has no effect (multiplies the odds by 1.0000...) and it has a large p-value so it is not statistically significant. Regarding genre, comedy, documentary, and short films show significantly higher odds of receiving a rating greater than 7, while drama films tend to have lower odds. However, the effect of the "Romance" and "Animation" genre does not appear statistically significant. All the predictors we used except the "Romance genre", and the "Animation genre" are statistically significant (at the 5% significance level) as their p-value is less than 0.5.

## Log odds and 95% Confidence Intervals

Note that the following work is done with the removal of the variable votes in order to make the model have a better fit. The final optimal model is found here @sec-final-refined-optimal-model
```{r}
#| echo: false
#confint(glm1) %>%
# kable()
```

```{r}
plot_model(glm1, show.values = TRUE,
           title = "Odds (Film rating higher than 7)", show.p = FALSE)
```

|                    | 2.5%    | 97.5%  |
|--------------------|---------|--------|
| (Intercept)        | -35.26  | -12.64 |
| year               | 0.00    | 0.02   |
| length             | -0.064  | -0.05  |
| budget             | 0.45    | 0.57   |
| genre: Animation   | -0.71   | 0.55   |
| genre: Comedy      | 2.77    | 3.47   |
| genre: Documentary | 4.79    | 6.53   |
| genre: Drama       | -2.04   | -1.10  |
| genre: Romance     | -126.17 | 5.33   |
| genre: Short       | 2.65    | 5.92   |

Thus from the plot and the values above our interpretations for the coefficients from before are confirmed. We interpret the odds ratios as follows: The lowest ones, romance genre films odds of getting a rating higher than 7 were 0.0004 times those of other genres, while the highest ones documentary films of getting a rating higher than 7 were 271.76 times higher than those of other genres. For each year increase in when the movie was released the odds of a rating higher than 7 increase by 1.01, while for each minute of increase of the movie's length the odds of a rating higher than 7 decreased (by a factor of 0.94). Finally, for a one million increase (\$) in budget the odds of a rating higher than 7 increased by a factor of 1.67.

```{r}
#Add log-odds to our data set
films <- films %>%
  mutate(logodds.great = predict(glm1))
#Add odds to our data set
films <- films %>%
  mutate(odds.great = exp(logodds.great))

#Add probabilities to our data set
films <- films %>%
  mutate(probs.great = fitted(glm1))


```

## Model Assessment and Model selection

```{r}
# ROC Curve Analysis
library(pROC)
roc_obj <- roc(films$GREAT, fitted(glm1))
plot(roc_obj)
auc(roc_obj)

```

From a first look at the ROC curve above and by the fact that the area under the curve is 0.95 which is close to 1, it seems that the model might be a good fit. Now to check whether the chi-squared approximation seems plausible to use for comparison with the deviance we must first check if the fitted values in the second column of the output with the exception of the first one are quite large (\>5).

```{r}
#| echo: false
p.hat <- predict(glm1 , type = "response")
fitted <- nrow(films) * p.hat

head(cbind(films$GREAT, round(fitted,2)))
#chi squared
qchisq(df = 2706 , p = 0.95)
```

So from the table above it can clearly be seen that they are indeed greater than 5. Thus, the chi-squared distribution as a measure of goodness of fit is valid here. So now we compare the residual deviance of the model ($1456.6$) to the 95th percentile of $\chi^2(2716-10) = 2828.132$ and it is obvious that $1456.6 < 2828.132$ . Therefore we don't have evidence of lack of fit.

We will now examine the use of the following model just to see if we can achieve a better fit with a better model:

$$
g(p_i) = \boldsymbol \Phi ^{-1}(p_i) = \mathbf{x}_i^\top \boldsymbol{\beta}
$$

where $\boldsymbol \Phi$ denotes the cumulative distribution function of the standard normal distribution.

```{r}
glm2 <- glm(GREAT ~ year + length + budget + genre , data = films , family= binomial(link = 'probit'))
summary(glm2)

```

From the output of the model above we are still able to conduct a goodness of fit test based on the residual deviance $(1473<2828.132)$ so no evidence of lack of fit here as well, but as the deviance of the logit model is slightly lower we can say that the fit of it is better. Finally a third choice of a link function we will consider is the complementary log-log link, with the model being:

$$
g(p_i) =  \log\left[-\log(1-p_i)\right] = \mathbf{x}_i^\top \boldsymbol{\beta}
$$

```{r}
glm3 <- glm(GREAT ~ year + length + budget + genre , data = films , family= binomial(link = 'cloglog'))
summary(glm3)
```

From the output from this third model above the residual deviance is $(1602<2828.132)$ so again there's no evidence of lack of fit. However it seems that the logit has the lowest of the deviances out of them so we would expect that it is indeed the better model out of the three.

```{r}

# Calculate AIC and BIC values
aic_values <- c(AIC(glm1), AIC(glm2), AIC(glm3))
bic_values <- c(BIC(glm1), BIC(glm2), BIC(glm3))


aic_bic_table <- data.frame(
  Model = c("logit", "probit", "cloglog"),
  AIC = aic_values,
  BIC = bic_values
)
aic_bic_table
```

Based on the table above it is clear that the logit model we chose was the most appropriate one out of the three.

## Probability plots based on logit model

```{r}
library(sjPlot)
plot_model(glm1,type="pred",terms=c("budget[all]"), axis.title=c("Film budget ($ in millions)", "Probability that the film is rated higher than 7"),
title="Probability of the film rated higher than 7 vs budget", ci.lvl=NA)

plot_model(glm1,type="pred",terms=c("genre[all]"), axis.title=c("Film genre", "Probability that the film is rated higher than 7"),
title="Probability of the film rated higher than 7 vs genre", ci.lvl=NA)


plot_model(glm1,type="pred",terms=c("length[all]"), axis.title=c("Film length (in minutes)", "Probability that the film is rated higher than 7"),
title="Probability of the film rated higher than 7 vs Film duration", ci.lvl=NA)
```

The probability plots above provide compelling insights into the influence of budget, genre, and length on the likelihood of a film receiving a rating greater than 7 on IMDb based on our logit model. Notably, budget emerges as a significant predictor, with each additional unit of financial investment substantially increasing the probability of achieving a higher rating. This finding underscores the importance of budget allocation in shaping audience perceptions and critical reception. Furthermore, genre analysis reveals distinct patterns: comedy, documentary, and short films exhibit notably favorable probabilities, suggesting a propensity for audience appeal and critical acclaim. Conversely, drama films present a contrasting scenario, indicating lower probabilities of receiving a higher rating. Finally, film length demonstrates a discernible trend, with shorter films showing slightly higher probabilities of attaining a rating above 7.
## Final refined optimal model {#sec-final-refined-optimal-model}

We will now remove all the genres with the high p-values that are statistically insignificant, and we will create a refined model with the following output:

```{r}
# Create a subset of the dataframe without rows containing "Romance" or "Animation" in the genre column
Refined_data <- subset(films, !(genre %in% c("Romance", "Animation")))

#Create the the final refined glm
refined_glm <-  glm(GREAT ~ year + length + budget + genre  , data = Refined_data , family= binomial(link = logit))
summary(refined_glm)
```

```{r}
#| echo: false
qchisq(df = 2506 , p = 0.95)
```

This model now is the best possible model while also satisfying the goodness of fit, as deviance is less than 2623.573 and we have no variables with p-value greater than 0.05. The equation for this model now has become:

$$
 \log\left(\frac{p_i}{1-p_i}\right) = -36.09 + 0.02 \cdot year -0.06 \cdot length + 0.54 \cdot budget + 3.20 \cdot  \boldsymbol{\mathbb{I}_{Com}} + 5.65 \cdot  \boldsymbol{\mathbb{I}_{Doc}} -1.54 \cdot  \boldsymbol{\mathbb{I}_{Dra}} +4.2 \cdot  \boldsymbol{\mathbb{I}_{Short}} 
$$

where $\boldsymbol {\mathbb{I}}$ is the indicator function for each genre, that is 0 if it is not the corresponding genre and 1 if it is. For example $\boldsymbol {\mathbb{I}}_{Com}$ is 1 if the film is a Comedy. Note that a film in our dataset can never be two genres.

```{r}
#| echo: true

#Odds ratios of variables
exp(0.016167)   #year
exp(-0.056306)  #length
exp(0.542263)   #budget
```

Based on this model, when all other variables are held constant, a 1% increase in the odds of the film getting a rating greater than 7 is observed for every 1 year increase in film release. In addition when all other variables are held constant, for each minute increase in the movie, there is a 6% decrease in the odds of the film getting a rating greater than 7, while for each million added to the movie's budget the odds of the film getting a rating higher than 7 increase by 72%. What's more from the coefficient estimates we are seeing again the the highest estimate is reached when the genre of the movie is a documentary, once again confirming our initial impressions.

```{r}
plot_model(refined_glm , show.values = TRUE, transform = NULL, title = "Log-Odds", show.p = FALSE)
```

**Example of a prediction:**

Here we will show an example of a prediction based on this refined model of a Comedy movie, released in 1982, with a duration of 100 minutes , budget of 10 million (\$).

```{r}
#| echo: true
#odds of having a rating higher than 7
exp(-36.085660 + (0.016167*1982) - (0.056306*100) + (0.542263*10) + 3.196425)  

#probability of having a rating higher than 7
exp(-36.085660 + (0.016167*1982) - (0.056306*100) + (0.542263*10) + 3.196425) / (1 + exp(-36.085660 + (0.016167*1982) - (0.056306*100) + (0.542263*10) + 3.196425))
```

Therefore, the odds of a film having a rating higher than 7 with these characteristics, is 0.34 and the probability of having a rating higher than 7 is approximately 26%.

# Conclusions {#sec-Conc}

NOTE THIS SHOULD BE ADDED TO EXPLORATORY ANALYSIS

```{r}

library(janitor)
films$GREAT <- as.factor(films$GREAT)
films$genre <- as.factor(films$genre)
levels(films$GREAT) <- c("less rating than 7" , "greater rating than 7")


films %>%
tabyl(genre, GREAT) %>%
adorn_percentages() %>%
adorn_pct_formatting() %>%
adorn_ns() 



```
